---
title: "Regression analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(choroplethrZip)
library(RSocrata)
```


```{r load, cache = TRUE}
with_detail = read.socrata("https://data.cityofnewyork.us/resource/ibte-hq4u.csv")
```

```{r tidy_data}
reg_prep = 
  with_detail %>% 
  select(incident_date_time, incident_type_desc, 
         property_use_desc, zip_code, borough_desc) %>% 
  #get a clean borough variable
  separate(borough_desc, sep = "- ", into = c("n", "borough")) %>% 
  select(-n) %>% 
  #make it easier to filter true fire alarms
  separate(incident_type_desc, into = c("incident_type", "incident_desc"), sep = "- ") %>%
  mutate(incident_type = str_sub(incident_type, 1, 3),
         incident_type = as.integer(incident_type)) %>% 
  #clean year and month
  mutate(month = as.factor(month.abb[as.integer(substr(incident_date_time, 6, 7))]),
         month = fct_relevel(month, "Jan", "Feb", "Mar", "Apr", "May", 
                             "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 
         year = str_sub(incident_date_time, 1, 4)) %>% 
  #get a season variable. I can't believe I'm using this ugly if_else
  mutate(season = 
           if_else(month %in% c("Dec", "Jan", "Feb"), "winter", 
                   if_else(month %in% c("Mar", "Apr", "May"), "spring", 
                           if_else(month %in% c("Jun", "Jul", "Aug"), "summer", "autumn"))), 
         season = as.factor(season)) %>% 
  #clean zip code, exclude unreasonable zip codes
  mutate(zip_code = str_sub(zip_code, 1, 5)) %>% 
  filter(zip_code != "" & zip_code != "99999" & 
         zip_code != "10691" & zip_code != "11251", 
         incident_type < 165, 
         year == "2013") %>% 
  group_by(zip_code, season) %>% 
  summarize(true_fire = n())

##pull population data from the choroplethrZip package
data("df_pop_zip")
pop_in_ny = 
  df_pop_zip %>% 
  rename(zip_code = region, 
         population = value)

reg_data = left_join(reg_prep, pop_in_ny, by = "zip_code")
```

We wanted to further explore the factors that affects the number of true fire cases in a specific area in 2013. We found that the population heat map was very similar to our fire cases distribution map, therefore we build a linear regression model to test factors that might have associations with the number of true fire cases. 

```{r reg}
fit = lm(true_fire ~ population + season, data = reg_data)
broom::tidy(fit) %>% 
  knitr::kable(digits = 3)
```


The model shows that population in an area is highly associated with the number of fire cases. In a given season, when average population increases by 1000, there will be one more true fire cases in an area. Also in summer there seems to be an average of 8 less true fire cases than in the autumn.  


If you are finding a place to live in New York City and you would like to live in a place that is least likely to caught on fire, we would suggest to live in a less populated area.  